---
title: "Homework 1"
author: "Giannopoulou Ilektra Eleni - 23202690"
date: today
format: pdf
fontsize: 13pt
---

# "Analysis of the Iris Dataset"

## Introduction

The \textcolor{blue}{iris} dataset provides the measurements in centimeters of the variables sepal length and width, and petal length and width, for 50 flowers from each of three species of iris. The species are ___Iris setosa___, ___Iris versicolor___, and ___Iris virginica___.


## Data Overview

The dataset consists of 150 observations with five variables: Sepal.Length, Sepal.Width, Petal.Length, Petal.Width, and Species. Each variable represents a measurement in centimeters, except for Species, which is a factor variable with three levels.

_Descriptive Statistics_:

```{r}
#| echo: false
#load the iris dataset
data(iris)

#generate summary statistics
summary_statistics = summary(iris[,1:4])

#print the table
knitr::kable(summary_statistics, caption = "Summary Statistics of Iris Dataset")
```


## Histogram of Sepal Width

_Approach Explanation_:

To analyze the \textcolor{blue}{iris} dataset, first of all we need to have loaded the data using the __data()__ function. Next, we will create a histogram of the _Sepal Width_ variable to visualize its distribution. To accomplish this task, we will use the __hist()__ function with custom colors and border parameters to improve the visual appeal of the plot.

```{r}
#| echo: false

#plotting histogram of sepal width
hist(iris$Sepal.Width,
     main = "Histogram of Sepal Width",
     xlab = "Sepal Width (cm)",
     col = "purple",
     border = "midnightblue")
```

This histogram illustrates the distribution of _Sepal Width_ across the \textcolor{blue}{iris} dataset, showing a __normal distribution__ with a central peak around the mean. The most frequent sepal width values range between about 2.5 and 3.5 cm confirming the result of the summary statistic above.



